Drug sensitivity prediction with normal inverse Gaussian shrinkage informed by external data

### 2020-04-22 docs/supplement.Rnw
commit before pull

% load packages
<<packages, include=FALSE, echo=FALSE>>=
library(cambridge)
@

% set options
<<settings, include=FALSE, echo=FALSE>>=
opts_knit$set(root.dir="..", base.dir="../figs")
opts_chunk$set(fig.align='center', fig.path="../figs/", 
               echo=FALSE, cache=FALSE, message=FALSE)
@

@@

\documentclass[a4paper,hidelinks]{article}
\usepackage{bm,amsmath,amssymb,amsthm,amsfonts,graphics,graphicx,epsfig,
rotating,caption,subcaption,natbib,appendix,titlesec,multicol,hyperref,verbatim,
bbm,algorithm,algpseudocode,pgfplotstable,threeparttable,booktabs,mathtools,
dsfont,parskip}

@@

% settings
\pgfplotsset{compat=1.13}
\setlength{\evensidemargin}{.5cm} \setlength{\oddsidemargin}{.5cm}
\setlength{\textwidth}{15cm}
\title{Supplementary material to: Drug efficacy prediction in cell lines}
\date{\today}
\author{Magnus M. M\"unch$^{1,2}$\footnote{Correspondence to: \href{mailto:m.munch@vumc.nl}{m.munch@vumc.nl}}, Mark A. van de Wiel$^{1,3}$, Sylvia Richardson$^{3}$, \\ and Gwena{\"e}l G. R. Leday$^{3}$}

@@

	\noindent
	1. Department of Epidemiology \& Biostatistics, Amsterdam Public Health research institute, VU University Medical Center, PO Box 7057, 1007 MB
	Amsterdam, The Netherlands\\
	2. Mathematical Institute, Leiden University, Leiden, The Netherlands \\
	3. MRC Biostatistics Unit, Cambridge Institute of Public Health, Cambridge, United Kingdom \\

	\section{Content overview}
	This document contains the Supplementary Material (SM) to the document `Drug efficacy prediction in cell lines'. In the following, this document is referred to as Main Document (MD).

	\section{Variational Bayes derivations}\label{sec:vbderivations}
	In the following all expectations are with respect to the variational posterior $Q$.
	\begin{align*}
	\log q_{\B} (\B) & \propto \E_{\mathbf{U}, \bsigma^2} [\log \mathcal{L}(\Y | \B, \U, \bsigma^2 )] + \E_{\bgamma^2, \bsigma^2} [\log \pi (\B | \bgamma^2, \bsigma^2)] \\
	& \propto - \frac{1}{2} \sum_{d=1}^D \sum_{i=1}^n \E_{\mathbf{u}_d, \sigma_d^2} \[ \frac{(y_{id} - \x_i \tr \bbeta_d - \mathbf{z}_i \tr \mathbf{u}_d)^2}{\sigma_d^2} \] - \frac{1}{2} \sum_{d=1}^D \sum_{j=1}^p \phi_{g(j)}^{-2} \E_{\gamma^2_d, \sigma^2_d} \( \frac{\beta_{jd}^2}{\gamma_d^2 \sigma_d^2} \), \\
	q_{\B} (\B) & \overset{D}{=} \prod_{d=1}^{D} \mathcal{N}_p (\bmu_d, \bSigma_d), \\
	& \text{ with } \bSigma_d = \E(\sigma_d^{-2})^{-1} [\X \tr \X + \E(\gamma_d^{-2}) \cdot \diag (\phi_{g(j)}^{-2}) ]^{-1}, \\
	& \text{ and } \bmu_d = [\X \tr \X + \E(\gamma_d^{-2}) \cdot \diag (\phi_{g(j)}^{-2}) ]^{-1} \X \tr [\y_d - \Z \cdot \E(\mathbf{u}_d)].
	\end{align*}

	\begin{align*}
	\log q_{\U}(\U) & \propto \E_{\mathbf{B}, \bsigma^2} [\log \mathcal{L}(\Y | \B, \U, \bsigma^2 )] + \E_{\mathbf{\Xi}_1, \dots, \mathbf{\Xi}_D, \bsigma^2} [\log \pi (\U | \mathbf{\Xi}_1, \dots, \mathbf{\Xi}_D, \bsigma^2)] \\
	& \propto - \frac{1}{2} \sum_{d=1}^D \sum_{i=1}^n \E_{\bbeta_d, \sigma_d^2} \[ \frac{(y_{id} - \x_i \tr \bbeta_d - \mathbf{z}_i \tr \mathbf{u}_d)^2}{\sigma_d^2} \] - \frac{1}{2} \sum_{d=1}^D \E_{\bm{\Xi}_d, \sigma_d^2}\( \frac{\mathbf{u}_d \tr \bm{\Xi}_d^{-1} \mathbf{u}_d}{\sigma_d^2} \), \\
	q_{\U}(\U) & \overset{D}{=} \prod_{d=1}^{D} \mathcal{N}_T (\mathbf{m}_d, \Sm_d), \\
	& \text{ with } \Sm_d = \E(\sigma_d^{-2})^{-1} [\Z \tr \Z + \E(\bm{\Xi}_d^{-1}) ]^{-1}, \\
	& \text{ and } \mathbf{m}_d = [\Z \tr \Z + \E(\bm{\Xi}_d^{-1}) ]^{-1} \Z \tr [\y_d - \X \cdot \E(\bbeta_d)].
	\end{align*}

	\begin{align*}
	\log q_{\bgamma^{2}}(\bgamma^{2}) & \propto \E_{\B, \bsigma^2} [\log \pi (\B | \bgamma^{2}, \sigma^2)] + \log \pi (\bgamma^{2}) \\
	& \propto -\frac{p}{2} \sum_{d=1}^D \log \gamma_d^{2} - \frac{1}{2} \sum_{d=1}^D \sum_{j=1}^p \phi_{g(j)}^{-2} \E \( \frac{\beta_{jd}^2}{\sigma^{2}_d}\) \gamma_d^{-2} - \frac{3}{2} \sum_{d=1}^D \log \gamma_d^{2} \\
	& \,\,\,\,\,\,\,\,\,\, - \frac{\lambda}{2 \theta^2} \sum_{d=1}^D (\gamma_d^2 - \theta)^2 \gamma_d^{-2} \\
	& \propto \sum_{d=1}^D \left\{ \( -\frac{p + 1}{2} - 1\) \log \gamma_d^{2} - \frac{\lambda}{2\theta^2} \gamma_d^2 - \frac{1}{2}\[ \lambda + \E(\sigma_d^{-2}) \sum_{j=1}^p \frac{\E(\beta_{jd}^2)}{\phi_{g(j)}^{2}}\] \gamma_d^{-2} \right\}, \\
	q_{\bgamma^{2}}(\bgamma^{2}) & \overset{D}{=} \prod_{d=1}^D \mathcal{GIG} \(-\frac{p+1}{2}, \frac{\lambda}{\theta_d^2}, \delta_{d} \), \\
	& \text{with } \delta_{d} =\E(\sigma_d^{-2}) \sum_{j=1}^p \frac{\E(\beta_{jd})^2 + \V(\beta_{jd})}{\phi_{g(j)}^2} + \lambda.
	\end{align*}

	\begin{align*}
	\log q_{\bsigma^{2}}(\bsigma^{2}) & \propto \E_{\B,\mathbf{U}} [\log \mathcal{L}(\Y | \B, \U, \bsigma^2 )] + \E_{\B, \bgamma^2} [\log \pi (\B | \bgamma^2, \bsigma^2)] \\
	& \,\,\,\,\,\,\,\,\,\, + \E_{\mathbf{U}, \bm{\Xi}_1, \dots, \bm{\Xi}_D} [\log \pi (\mathbf{U} | \bm{\Xi}_1, \dots, \bm{\Xi}_D, \bsigma^2)] + \log \pi (\bsigma^{2}) \\
	& \propto -\frac{n}{2} \sum_{d=1}^D \log \sigma_d^{2} - \frac{1}{2} \sum_{d=1}^D \sum_{i=1}^n \E [(y_{id} - \x_i \tr \bbeta_d - \mathbf{z}_i \tr \mathbf{u}_d)^2] \sigma_d^{2} - \frac{p}{2} \sum_{d=1}^D \log \sigma_d^{-2} \\
	& \,\,\,\,\,\,\,\,\,\, - \frac{1}{2} \sum_{d=1}^D \sum_{j=1}^p \E( \gamma_d^{-2}) \frac{\E ( \beta_{jd}^2)}{\phi_{g(j)}^{2}} \sigma_d^{-2} - \frac{T}{2} \sum_{d=1}^D \log \sigma_d^{2} - \frac{1}{2} \sum_{d=1}^D \E(\mathbf{u}_d \tr \bm{\Xi}_d^{-1} \mathbf{u}) \sigma_d^{-2} \\
	& \,\,\,\,\,\,\,\,\,\, - \frac{3}{2} \sum_{d=1}^D \log \sigma^2_d \\
	& = -\sum_{d=1}^{D} \(\frac{n + p + T + 3}{2} \) \log \sigma_d^{2} - \sum_{d=1}^{D} \frac{1}{2}\Bigg\{ \sum_{i=1}^n \E [(y_{id} - \x_i \tr \bbeta_d - \mathbf{z}_i \tr \mathbf{u}_d)^2] \\
	& \,\,\,\,\,\,\,\,\,\, + \E(\gamma_d^{-2})\sum_{j=1}^p \frac{\E(\beta_{jd}^2)}{\phi_{g(j)}^2} + \E(\mathbf{u}_{d} \tr \bm{\Xi}_d^{-1} \mathbf{u}_d) \Bigg\} \sigma_d^{-2}, \\
	q_{\bsigma^{2}}(\bsigma^{2}) & \overset{D}{=} \prod_{d=1}^D \Gamma^{-1} \(\frac{n + p + T + 1}{2}, \zeta_{d} \), \\
	& \text{ with } \zeta_{d} = \frac{\y_d \tr \y_d}{2} - \y_d \tr [\X \E(\bbeta_d) + \Z \E(\mathbf{u}_d)] + \E(\bbeta_d \tr) \X \tr \Z \E (\mathbf{u}_d) \\
	& \,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\, + \frac{\trace [\X \tr \X \V(\bbeta_d)]}{2} + \frac{\E(\bbeta_d \tr) \X \tr \X \E(\bbeta_d)}{2} + \frac{\trace [\Z \tr \Z \V(\mathbf{u}_d)]}{2} \\
	& \,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\, + \frac{\E(\mathbf{u}_d \tr) \Z \tr \Z \E(\mathbf{u}_d)}{2} + \frac{\E(\gamma_d^{-2})}{2} \sum_{j=1}^p \frac{\E(\beta_{jd})^2 + \V(\beta_{jd})}{\phi_{g(j)}^2} \\
	& \,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,+ \frac{\E(\mathbf{u}_d \tr) \E(\bm{\Xi}_d^{-1}) \E(\mathbf{u}_d)}{2} + \frac{\trace [\E(\bm{\Xi}_d^{-1}) \V(\mathbf{u}_d)]}{2},
	\end{align*}

	\begin{align*}
	\log q_{\bm{\Xi}_1, \dots, \bm{\Xi}_D}(\bm{\Xi}_1, \dots, \bm{\Xi}_D) & \propto \E_{\mathbf{U}, \bm{\sigma}^2} [\log \pi (\mathbf{U} | \bm{\Xi}_1, \dots, \bm{\Xi}_D, \bsigma^2)] + \log \pi (\bm{\Xi}_1, \dots, \bm{\Xi}_D) \\
	& \propto -\frac{1}{2} \sum_{d=1}^D \log |\bm{\Xi}_d| - \frac{1}{2} \sum_{d=1}^D \E \(\frac{\mathbf{u}_d \tr \bm{\Xi}_d^{-1} \mathbf{u}_d}{\sigma_d^2} \) \\
	& \,\,\,\,\,\,\,\,\,\, - \frac{k + T + 1}{2} \sum_{d=1}^D \log |\bm{\Xi}_d| - \frac{1}{2} \sum_{d=1}^D \trace (\bm{\Omega} \bm{\Xi}_d^{-1}) \\
	& = -\frac{\tau + T + 2}{2} \sum_{d=1}^D \log |\bm{\Xi}_d| - \frac{1}{2} \sum_{d=1}^D \trace \left\{ \[\E(\sigma_d^{-2}) \E( \mathbf{u}_d \mathbf{u}_d \tr) + \bm{\Omega}\] \bm{\Xi}_d^{-1} \right\}, \\
	q_{\bm{\Xi}_1, \dots, \bm{\Xi}_D}(\bm{\Xi}_1, \dots, \bm{\Xi}_D) & \overset{D}{\propto} \prod_{d=1}^D \mathcal{W}_T^{-1} (\bm{\Psi}_d, \nu + 1),  \\ 
	& \text{ with } \bm{\Psi}_d =\E(\sigma^{-2}_d) \[\V(\mathbf{u}_d) + \E(\mathbf{u}_d) \E( \mathbf{u}_d \tr) \] + \bm{\Omega}.
	\end{align*}

	\section{Simulations}
	\subsection{Models}
	In order to investigate the inverse Gaussian prior model and its estimation, we consider a simpler model. The simpler model ignores tissue effects and fixes $\forall g: \phi_g^2=1$:
	\begin{align*}
  y_{id} | \bm{\beta}_d, \sigma_d^2 & \sim \mathcal{N} (\x_i \tr \bm{\beta}_d, \sigma_d^2), \\
  \beta_{jd} | \gamma_d^2, \sigma_d^2 & \sim p(\beta_{jd}), \\
  \gamma_d^2 & \sim p(\gamma_d^2), \\
  \sigma_d^2 & \sim 1/\sigma_d^3.
  \end{align*}

  \cite{moran_variance_2018} argue that in regression models, $\bm{\beta}_d$ and $\sigma_d^2$ should be independent \textit{a priori}. However, in our case that is debatable. The $D$ regressions need not be on the same scale, so some form of calibration of the $\bm{\beta}_d$ posteriors to this scale is required. A simple way of doing this is to include the scales $\sigma_d^2$ in the priors for the $\bm{\beta}_d$. We call the two versions of the model conjugate (with $\bm{\beta}_d$ and $\sigma_d^2$ dependent):
  $$
  \beta_{jd} | \gamma_d^2, \sigma_d^2 \sim \mathcal{N} (0, \sigma_d^2 \gamma_d^2)
  $$
  and non-conjugate (with $\bm{\beta}_d$ and $\sigma_d^2$ independent):
  $$
  \beta_{jd} | \gamma_d^2, \sigma_d^2 \sim \beta_{jd} | \gamma_d^2 \sim \mathcal{N} (0, \gamma_d^2),
  $$
  and compare them in our simulations.

	In addition to the inverse Gaussian model for the prior variances $\gamma_d^2$ (introduced in the MD):
	$$
  \gamma_d^2 \sim \mathcal{IG}(\theta_d, \lambda_d),
  $$
	we investigate a generalisation of the models in \cite{kpogbezan_empirical_2017} and \cite{leday_gene_2017}, where we draw the $\gamma_d^2$ from an inverse Gamma distribution; :
	$$
	\gamma_d^2 \sim \Gamma^{-1}(\eta_d/2, \lambda_d/2).
	$$
	For the sake of comparability, we have rescaled the inverse Gamma parameters and reparametrised both distributions as generalized inverse Gaussians:
	\begin{align*}
  \gamma_d^2 & \sim \mathcal{GIG}(-1/2, \lambda_d \theta_d^{-2}, \lambda_d), \\
  \gamma_d^2 & \sim \mathcal{GIG}(-\eta_d/2, 0, \lambda_d).
  \end{align*}
	The two models coincide if $\theta_d \to \infty$ and $\eta_d = 1$.

>>>>

% set options
<<settings, include=FALSE, echo=FALSE>>=
opts_knit$set(root.dir="..", base.dir="../figs")
opts_chunk$set(fig.align='center', fig.path="../figs/", 
               echo=FALSE, cache=FALSE, message=FALSE, fig.pos='!ht')
knit_hooks$set(document=function(x) {
  sub('\\usepackage[]{color}', '\\usepackage{xcolor}', x, fixed=TRUE)})
options(knitr.kable.NA="", digits=3)
@

@@

\documentclass[a4paper,hidelinks]{article}
\usepackage{bm,amsmath,amssymb,amsthm,amsfonts,graphics,graphicx,epsfig,
rotating,caption,subcaption,natbib,appendix,titlesec,multicol,hyperref,verbatim,
bbm,algorithm,algpseudocode,pgfplotstable,threeparttable,booktabs,mathtools,
dsfont,parskip,xr}
\externaldocument[md-]{manuscript}

@@

% settings
\graphicspath{{../figs/}}
\pgfplotsset{compat=1.16}
\setlength{\evensidemargin}{.5cm} \setlength{\oddsidemargin}{.5cm}
\setlength{\textwidth}{15cm}
\title{Drug sensitivity prediction with normal inverse Gaussian shrinkage 
informed by external data}
\date{\today}
\author{Magnus M. M\"unch$^{1,2}$\footnote{Correspondence to: 
\href{mailto:m.munch@amsterdamumc.nl}{m.munch@amsterdamumc.nl}}, Mark A. van de 
Wiel$^{1,3}$, Sylvia Richardson$^{3}$, \\ and Gwena{\"e}l G. R. Leday$^{3}$}

@@

	\noindent
	1. Department of Epidemiology \& Biostatistics, Amsterdam UMC, VU University, 
	PO Box 7057, 1007 MB Amsterdam, The Netherlands \\
	2. Mathematical Institute, Leiden University, Leiden, The Netherlands \\
	3. MRC Biostatistics Unit, Cambridge Institute of Public Health, Cambridge,
	United Kingdom

	\section{Content overview}
	This document contains the Supplementary Material (SM) to the document 
	`Drug sensitivity prediction with the normal inverse Gaussian prior 
	informed by external data'. In the 
	following, this document is referred to as Main Document (MD).

	Section \ref{sec:prior} describes the moments of the normal
	inverse Gaussian (NIG) prior, Section \ref{sec:vbderivations} gives
	the variational Bayes (BV) derivations that lead to the estimating equations 
	in the MD, Section \ref{sec:elbo} gives the evidence lower bound (ELBO) that 
	is maximised during the VB iterations. In Section \ref{sec:computation} 
	efficient VB calculations are given, while Section 
	\ref{sec:ratiosmodifiedbessels} outlines a computationally stable method
	to calculate ratios of modified Bessel functions. Section \ref{sec:eb} shows
	that the empirical Bayes (EB) scale estimates $\bm{\lambda}$ are non-negative.
	The last Section \ref{sec:gibbssampler} introduces a Gibbs sampling scheme
	to sample from the posterior corresponding to the NIG model.

	\section{Normal inverse Gaussian prior moments}\label{sec:prior}
	Let $\beta \sim \mathcal{N}(0,\sigma^2 \tau^2 \gamma^2)$, where 
	$\sigma^2$ and $\tau^2$ are fixed and $\gamma^2 \sim 
	\mathcal{IG}(\phi,\lambda_{\text{feat}})$, then
	\begin{align*}
	  \E(\beta | \sigma^2, \tau^2) & = 0, \\
	  \V(\beta | \sigma^2, \tau^2) & = \sigma^2 \tau^2 \phi, \\
	  \mathcal{K}(\beta | \sigma^2, \tau^2) & = 
	  \frac{3\phi}{\lambda_{\text{feat}}} + 3,
	\end{align*}
	where $\mathcal{K}(\cdot)$ denotes the kurtosis. Now letting $\tau^2
	\sim \mathcal{IG}(\chi,\lambda_{\text{drug}})$ and integrating it out we have:
	\begin{align*}
	  \E(\beta | \sigma^2) & = 0, \\
	  \V(\beta | \sigma^2) & = \sigma^2 \chi \phi, \\
	  \mathcal{K}(\beta | \sigma^2) & = 
	  3\(\frac{\phi}{\lambda_{\text{feat}}} \frac{\chi}{\lambda_{\text{drug}}} +
	  \frac{\phi}{\lambda_{\text{feat}}} + \frac{\chi}{\lambda_{\text{drug}}}\) + 
	  3.
	\end{align*}

	\section{Hyperparameter settings}\label{sec:hyperparameters}
	For the NIG shrinkage weights prior in Figure \ref{md-fig:dens_kappa}a the 
	following hyperparameter settings (in terms of the inverse Gaussian prior) 
	were used: $\lambda=1$, $\phi=10$ (solid),
	$\lambda=0.1$, $\phi=10$ (dashed), and $\lambda=1$, $\phi=1$ (dotted).
	The Student`s $t$ prior on the regression parameters $\beta_{jd}$ arises if
	$\gamma_{jd}^2 \sim \Gamma^{-1}(a,b)$, with $\Gamma^{-1}(a,b)$ an inverse
	Gamma distribution with shape $a$ and scale $b$. Figure 
	\ref{md-fig:dens_kappa}b is then obtained by setting: $a=0.9$, $b=1.3$ 
	(solid), $a=1$, $b=4$ (dashed), and $a=1$, $b=0.2$ (dotted).
	The lasso prior results from an exponential prior on $\gamma_{jd}^2$. To
	create Figure \ref{md-fig:dens_kappa}c the following settings were
	used for exponential rate $\lambda$: $\lambda=1$ (solid), $\lambda=0.2$ 
	(dashed), and $\lambda=0.1$ (dotted).

	\section{Jeffrey's prior}\label{sec:jeffreysprior}
	We parametrise the data with a Gaussian distribution with unknown mean and 
	variance, i.e., $y_i \sim \mathcal{N} (\mu, \sigma^2)$. This results in the
	following Hessian (matrix of second derivatives) of the log density:
	$$
	\mathbf{H}(\log f(y_i | \mu, \sigma^2)) = \begin{bmatrix}
	-\frac{1}{\sigma^2} & y_i - \mu \\
	y_i - \mu & \frac{1}{2\sigma^4} - \frac{(y_i - \mu)^2}{\sigma^6}.
	\end{bmatrix}
	$$
	Jeffreys prior \cite[]{jeffreys_invariant_1946} is now found as:
	$$
	\pi(\mu, \sigma^2) \propto |\mathcal{I}(\mu, \sigma^2)|^{1/2} = 
	|-\E (\mathbf{H})|^{1/2} = \begin{vmatrix}
	\frac{1}{\sigma^2} & 0 \\
	0 & \frac{1}{2\sigma^4}
	\end{vmatrix}^{1/2} \propto 1/\sigma^3,
	$$
	where $\mathcal{I}(\theta)$ denotes the Fisher information of $\theta$.

	\section{Variational Bayes derivations}\label{sec:vbderivations}
	For clarity we have indexed the variational 
	density functions with their respective parameters, which we omitted in the 
	MD. In the following all expectations are with respect to the 
	variational posterior $Q_d$. The variational posterior for $\bbeta_d$ is 
	found as follows:
	\begin{align*}
	  \log q_{\bbeta_d} (\bbeta_d) & \propto \E [\log \mathcal{L}
	  (\y_d | \bbeta_d, \sigma_d^2 )] + \E [\log \pi (\bbeta_d | \bgamma_{d}^2, 
	  \sigma_d^2, \tau_d^2)] \\
	  & \propto - \frac{1}{2} \sum_{i=1}^n 
	  \E \[ \frac{(y_{id} - \x_i \tr \bbeta_d)^2}{\sigma_d^2} \] - 
	  \frac{1}{2} \sum_{j=1}^p 
	  \E \( \frac{\beta_{jd}^2}{\gamma_{jd}^2 \sigma_d^2 \tau_d^2} \), \\
	  q_{\bbeta_d} (\bbeta_d) & \overset{D}{=} 
	  \mathcal{N}_p (\bmu_d, \bSigma_d), \\
	  & \text{with } \bSigma_d = \E(\sigma_d^{-2})^{-1} 
	  \{\X \tr \X + \E(\tau_d^{-2})\diag [\E(\gamma_{jd}^{-2})] \}^{-1}, \\
	  & \text{and } \bmu_d = \{\X \tr \X + \E(\tau_d^{-2})
	  \diag [\E(\gamma_{jd}^{-2})] \}^{-1} 
	  \X \tr \y_d.
	\end{align*}
	The variational posterior for $\gamma_{jd}^2$ is given by:
	\begin{align*}
	  \log q_{\gamma_{jd}^{2}}(\gamma_{jd}^{2}) & \propto \E 
	  [\log \pi (\beta_{jd} | \gamma_{jd}^{2}, \sigma_d^2, \tau_d^2)] + 
	  \log \pi(\gamma_{jd}^{2}) \\
	  & \propto -\frac{1}{2} \log \gamma_{jd}^{2} - \frac{1}{2} 
	  \E \( \frac{\beta_{jd}^2}{\sigma^{2}_d \tau_d^2}\) \gamma_{jd}^{-2} - 
	  \frac{3}{2} \log \gamma_{jd}^{2} - 
	  \frac{\lambda_{\text{feat}} \{\gamma_{jd}^2 - 
	  \phi_{jd}\}^2}
	  {2\phi_{jd}^{2}} \gamma_{jd}^{-2} \\
	  & \propto \( -1 - 1\) \log \gamma_{jd}^{2} - 
	  \frac{\lambda_{\text{feat}}}{2\phi_{jd}^2} \gamma_{jd}^2 - 
	  \frac{1}{2}\[ \lambda_{\text{feat}} + \E(\tau_d^{-2})\E(\sigma_d^{-2}) 
	  \E(\beta_{jd}^2)\] \gamma_{jd}^{-2} , \\
	  q_{\gamma_{jd}^{2}}(\gamma_{jd}^{2}) & \overset{D}{=} \mathcal{GIG} 
	  \(-1, \lambda_{\text{feat}}/\phi_{jd}^{2}, \delta_{jd} \), \\
	  & \text{with } \delta_{jd} =\E(\tau_d^{-2})\E(\sigma_d^{-2})
	  \[\E(\beta_{jd})^2 + \V(\beta_{jd})\] + \lambda_{\text{feat}}.
	\end{align*}
	Similarly, we derive the variational posterior for $\tau_d^2$ as:
	\begin{align*}
	  \log q_{\tau_{d}^{2}}(\tau_{d}^{2}) & \propto \E 
	  [\log \pi (\bbeta_{d} | \bgamma_{d}^{2}, \sigma_d^2, \tau_d^2)] + 
	  \log \pi(\tau_{d}^{2}) \\
	  & \propto -\frac{p}{2} \log \tau_{d}^{2} - \frac{1}{2} \sum_{j=1}^p
	  \E \( \frac{\beta_{jd}^2}{\sigma^{2}_d \gamma_{jd}^{2}}\)\tau_d^{-2}  - 
	  \frac{3}{2} \log \tau_{d}^{2} - 
	  \frac{\lambda_{\text{drug}} \{\tau_{d}^2 - \chi_{d}\}^2}
	  {2\chi_{d}^{2}} \tau_{d}^{-2} \\
	  & \propto \( -\frac{p + 1}{2} - 1\) \log \tau_{d}^{2} - 
	  \frac{\lambda_{\text{drug}}}{2\chi_{d}^2} \tau_{d}^2 - 
	  \frac{1}{2}\[ \lambda_{\text{drug}} + \E(\sigma_d^{-2}) \sum_{j=1}^p 
	  \E(\gamma_{jd}^{-2}) \E(\beta_{jd}^2)\] \tau_{d}^{-2} , \\
	  q_{\tau_{d}^{2}}(\tau_{d}^{2}) & \overset{D}{=} \mathcal{GIG} 
	  \(-\frac{p + 1}{2}, \lambda_{\text{drug}}/\chi_{d}^{2}, \eta_{d} \), \\
	  & \text{with } \eta_{d} =\E(\sigma_d^{-2}) \sum_{j=1}^p
	  \E(\gamma_{jd}^{-2})\[\E(\beta_{jd})^2 + \V(\beta_{jd})\] + 
	  \lambda_{\text{drug}}.
	\end{align*}
	Lastly, we have the variational posterior for $\sigma_d^2$:
	\begin{align*}
	  \log q_{\sigma_d^{2}}(\sigma_d^{2}) & \propto \E 
	  [\log \mathcal{L}(\y_d | \bbeta_d, \sigma_d^2 )] + 
	  \E [\log \pi (\bbeta_d | \bgamma_d^2, \sigma_d^2,\tau_d^2)] + 
	  \log \pi (\sigma_d^{2}) \\
	  & \propto -\frac{n}{2} \log \sigma_d^{2} - \frac{1}{2} \sum_{i=1}^n 
	  \frac{\E [(y_{id} - \x_i \tr \bbeta_d)^2]}{\sigma_d^{2}} - 
	  \frac{p}{2} \log \sigma_d^{2} - \frac{1}{2} \sum_{j=1}^p 
	  \E \( \frac{\beta_{jd}^2}{\gamma_{jd}^{2}\tau_d^2}\) \sigma_d^{-2} - 
	  \frac{3}{2} \log \sigma^2_d \\
	  & = \(-\frac{n + p + 1}{2} - 1 \) \log \sigma_d^{2} \\
	  & \,\,\,\,\,\,\,\,\,\, - 
	  \frac{1}{2}\left\{ \sum_{i=1}^n \E [(y_{id} - \x_i \tr \bbeta_d)^2] + 
	  \E(\tau_d^{-2}) \sum_{j=1}^p \E(\gamma_{jd}^{-2})
	  \E(\beta_{jd}^2) \right\} \sigma_d^{-2}, \\
	  q_{\sigma_d^{2}}(\sigma_d^{2}) & \overset{D}{=} 
	  \Gamma^{-1} \(\frac{n + p + 1}{2}, \zeta_{d} \), \\
	  & \text{with } \zeta_{d} = \frac{1}{2} \bigg\{\mathbf{y}_d \tr 
	  \mathbf{y}_d - 2 \mathbf{y}_d \tr \X \E(\bbeta_d) + 
    \trace [ \X \tr \X \V(\bbeta_d)] + 
    \E(\bbeta_d \tr)  \X \tr \X \E(\bbeta_d ) \\
    & \,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\, + 
    \E(\tau_d^{-2})\trace \{ \diag[\E(\gamma_{jd}^{-2})] \V(\bbeta_d)\} + 
    \E(\tau_d^{-2})\E(\bbeta_d \tr) \diag[\E(\gamma_{jd}^{-2})] \E(\bbeta_d )
    \bigg\}.
  \end{align*}
  If we fill in the expectations and variances, we arrive at the estimating
  equations in MD equations (\ref{md-eq:vbequations}).

  \section{Evidence lower bound}\label{sec:elbo}
  The evidence lower bound that is maximised during the iterations is: 
  $\text{ELBO}^{(l)}=\sum_{d=1}^D \text{ELBO}_d^{(l)}$, with:
  \begin{align*}
    \text{ELBO}&^{(l)}_d \\
    & = -\frac{n+p+1}{2} \log \pi + \frac{p+1-n}{2}\log 2 + \frac{n+1}{2} + p
    + \log \Gamma \(\frac{n + p + 1}{2} \) \\
    & - \frac{n + p + 1}{2} \log \zeta_d^{(l)} + 
    \frac{1}{2} \log |\bSigma_d^{(l)}| - 
    \frac{n+p+1}{4\zeta_d^{(l)}}  \bigg[\mathbf{y}_d \tr \mathbf{y}_d -
    2 \mathbf{y}_d \tr \X \bm{\mu}_d^{(l)} + 
    \trace ( \X \tr \X \bm{\Sigma}_d^{(l)}) \\ 
    & + (\bm{\mu}_d^{(l)}) \tr \X \tr \X \bm{\mu}_d^{(l)} +
    g_{d}^{(l)} \trace \[ \diag (b_{jd}^{(l)}) \bm{\Sigma}_d^{(l)}\] + 
    g_{d}^{(l)} (\bm{\mu}_d^{(l)}) \tr \diag (b_{jd}^{(l)}) 
    \bm{\mu}_d^{(l)}\bigg] + p \log \lambda_{\text{feat}}^{(l)} \\
    & + \frac{p + 3}{4} \log \lambda_{\text{drug}}^{(l)} + 
    \frac{\eta_d^{(l)} - \lambda_{\text{drug}}^{(l)}}{2} g_d^{(l)} + 
    \sum_{j=1}^p \frac{\delta_{jd}^{(l)} - \lambda_{\text{feat}}^{(l)}}{2} 
    b_{jd}^{(l)} + \frac{\lambda_{\text{drug}}^{(l)}}{\chi_d^{(l)}} +
    \sum_{j=1}^p \frac{\lambda_{\text{feat}}^{(l)}}{\phi_{jd}^{(l)}} + \\
    & - \frac{p + 1}{2} \log \chi_d^{(l)} - \sum_{j=1}^p \log \phi_{jd}^{(l)}
    -\frac{p+1}{4} \log \eta_d^{(l)} - 
    \frac{1}{2} \sum_{j=1}^p \log \delta_{jd}^{(l)} \\
    & + \log K_{\frac{p+1}{2}}\(\sqrt{\eta_{d}^{(l)}
    \lambda_{\text{drug}}^{(l)}}/
    \chi_{d}^{(l)}\) + \sum_{j=1}^p \log K_{1}\(\sqrt{\delta_{jd}^{(l)}
    \lambda_{\text{feat}}^{(l)}}/\phi_{jd}^{(l)}\).
  \end{align*}

  \section{Efficient computation}\label{sec:computation}
  The empirical Bayes updates require the quantities
  $e_{jd}$, $b_{jd}$, $g_d$, and $f_d$, which in turn depend on the quantities 
  $\diag(\bSigma_d)$, $\bmu_{d}$, $\y_d \tr \X \bmu_d$, 
  $\trace(\X \tr \X \bSigma_d)$, $\bmu_d \tr \X \tr \X \bmu_d$,
  $g_d\trace[\diag(b_{jd})\bSigma_d]$, and $g_d\bmu_d \tr \diag(b_{jd}) \bmu_d$.
  Let $\mathbf{H}_d=g_d \diag[(b_{jd}^{(h)})^{-1}]$. Then, the first two
  quantities are efficiently calculated as:
  \begin{subequations}\label{eq:complexity1}
    \begin{align}
      \diag (\bm{\Sigma}_d^{(h+1)}) & = (a_d^{(h)})^{-1} 
      \H_d^{(h)} - (a_d^{(h)})^{-1} \{ [ \H_d^{(h)} \X \tr 
      ( \I_n + \X \H_d^{(h)} \X \tr )^{-1} ] \circ
      (\H_d^{(h)} \X \tr ) \} \mathbf{1}_{n \times 1}, \\
      \bmu_d^{(h+1)}) & = \H_d^{(h)} \X \tr \y_d - \H_d^{(h)} \X \tr 
      ( \I_n + \X \H_d^{(h)} \X \tr )^{-1} \X \H_d^{(h)} \X \tr \y_d.
    \end{align}
  \end{subequations}
  The quantities $\y_d \tr \X \bmu_d$, $\bmu_d \tr \X \tr \X \bmu_d$,
  $g_d\trace[\diag(b_{jd})\bSigma_d]$, and $g_d\bmu_d \tr \diag(b_{jd}) \bmu_d$
  are easily calculated from the first two. For the remaining one we have:
  \begin{align}\label{eq:complexity2}
    \trace(\X \tr \X \bSigma_d^{(h+1)})  = 
    & (a_d^{(h)})^{-1} \mathbf{1}_{1 \times p} \H_d^{(h)} \X \tr 
    \mathbf{1}_{n \times 1} \nonumber \\  
    & - (a_d^{(h)})^{-1} \mathbf{1}_{1 \times p} \{ [  \H_d^{(h)} \X \tr 
    ( \I_n + \X \H_d^{(h)} \X \tr )^{-1} \X \H_d^{(h)} \X \tr ] \circ 
    \X \tr \} \mathbf{1}_{n \times 1}.
  \end{align}
  Both (\ref{eq:complexity1}) and (\ref{eq:complexity2}) are operations of
  complexity $\mathcal{O}(pn^2)$ instead of $\mathcal{O}(p^3)$ for naive
  calculation.
